---
title: "hw 2"
output: word_document
---

```{r}
data(iris)
head(iris)

sp_ids = unique(iris$Species)

output = matrix(0, nrow=length(sp_ids), ncol=ncol(iris)-1)
rownames(output) = sp_ids
colnames(output) = names(iris[ , -ncol(iris)])

for(i in seq_along(sp_ids)) {
    iris_sp = subset(iris, subset=Species == sp_ids[i], select=-Species)
    for(j in 1:(ncol(iris_sp))) {
        x = 0
        y = 0
        if (nrow(iris_sp) > 0) {
            for(k in 1:nrow(iris_sp)) {
                x = x + iris_sp[k, j]
                y = y + 1
            }
            output[i, j] = x / y 
        }
    }
}
output

```
1. 
Output is mean value of each characteristic for each species.

2. 
Loop through species names, form subset for each species
   Loop through each column in subset
   If row number is greater than zero
       loop through column and calculate mean of column

3. X could be renamed sumcharacteristic, y could be renamed ntotal, and output could be renamed averagecharacteristic or meancharacteristic.

4.
Using a mean function instead of the third nested for loop would be more efficient.

5. 
```{r}
y = 0
x = 1:10

for(i in x){
    y[i] = sum(x[1:i])
    print(y)
}
```

6.
```{r}
y = 0
x = 1:10

for(i in x){
    y[i] = sum(x[1:i])
    if(y[i]<=10){
      print(y)
    }
    else{
      print(NA)
    }
}
```

7.
```{r}

sum_1 = function(x, cutoff=10) {
    output = NULL
    for(i in seq_along(x)) {
    output[i] = sum(x[1:i])
        if(output[i] > cutoff) {
            output[i] = NA
        }
    }
    return(output)
}
sum_1(x)
```
