---
title: "spatial modeling assignment"
author: "Merry Andersen"
date: "February 17, 2016"
output: word_document
---
```{r}
library(vegan)
library(nlme)
data(BCI)
?BCI
dim(BCI)
## UTM Coordinates (in metres)
BCI_xy = data.frame(x = rep(seq(625754, 626654, by=100), each=5), 
                    y = rep(seq(1011569,  1011969, by=100), len=50))
sp_ids = c("Cordia.lasiocalyx", "Hirtella.triandra",
           "Picramnia.latifolia", "Quassia.amara",
           "Tabernaemontana.arborea", "Trattinnickia.aspera", 
           "Xylopia.macrantha", "Drypetes.standleyi")

spsubset = subset(BCI, select=sp_ids)
head(spsubset)
BCI_sub = data.frame(BCI_xy, spsubset)
head(BCI_sub)

```

Model 1: general least squares with Trattinickia aspera as the predictive variable and Drypetes standleyi as the response variable
```{r}

#gls ~ Trattinnickia.aspera
modtratt = gls(Drypetes.standleyi ~ Trattinnickia.aspera, data=BCI_sub)
summary(modtratt)
plot(Variogram(modtratt, form= ~ x + y))


#modtratt with exponential error
modtratt1 = gls(Drypetes.standleyi ~ Trattinnickia.aspera, data= BCI_sub, corr= corExp(form= ~ x + y))
summary(modtratt1)
plot(Variogram(modtratt1, form= ~ x + y))


#modtratt with exponential error and nugget
modtratt1nug = gls(Drypetes.standleyi ~ Trattinnickia.aspera, data= BCI_sub, corr= corExp(form= ~ x + y, nugget = T))
summary(modtratt1nug)
plot(Variogram(modtratt1nug, form= ~ x + y))


#anova of models
anova(modtratt, modtratt1, modtratt1nug, test = F)

```
Both models show strong correlation between D. standleyi and T. aspera. Including the exponential error decreased the regression coefficient and the p-value for T. aspera. In the ANOVA the AIC and BIC were reduced by including the error, indicating a better fit for the model, but the graph shows an exponential trend along the data points. This means a different regression model is probably needed. Adding a nugget did not improve the model fit.

Model 2: D. standleyi as response variable, all other species as predictive variables
```{r}

#gls ~ all species
modall = gls(Drypetes.standleyi ~ Cordia.lasiocalyx + Hirtella.triandra + Picramnia.latifolia + Quassia.amara + Tabernaemontana.arborea + Trattinnickia.aspera + Xylopia.macrantha, data=BCI_sub)
summary(modall)
plot(Variogram(modall, form= ~ x + y))


#modall with exponential error
modall1 = gls(Drypetes.standleyi ~ Cordia.lasiocalyx + Hirtella.triandra + Picramnia.latifolia + Quassia.amara + Tabernaemontana.arborea + Trattinnickia.aspera + Xylopia.macrantha, data= BCI_sub, corr= corExp(form= ~ x + y))
summary(modall1)
plot(Variogram(modall1, form= ~ x + y))

#try exponential error with nugget
modall1nug = gls(Drypetes.standleyi ~ Cordia.lasiocalyx + Hirtella.triandra + Picramnia.latifolia + Quassia.amara + Tabernaemontana.arborea + Trattinnickia.aspera + Xylopia.macrantha, data= BCI_sub, corr= corExp(form= ~ x + y, nugget = T))
summary(modall1nug)
plot(Variogram(modall1nug, form= ~ x + y))


#There seems to be systematic error with the exponential correlation so try Gaussian instead
modallG = gls(Drypetes.standleyi ~ Cordia.lasiocalyx + Hirtella.triandra + Picramnia.latifolia + Quassia.amara + Tabernaemontana.arborea + Trattinnickia.aspera + Xylopia.macrantha, data= BCI_sub, corr= corGaus(form= ~ x + y))
plot(Variogram(modallG, form= ~ x + y))

#Try Gaussian with nugget
modallGnug = gls(Drypetes.standleyi ~ Cordia.lasiocalyx + Hirtella.triandra + Picramnia.latifolia + Quassia.amara + Tabernaemontana.arborea + Trattinnickia.aspera + Xylopia.macrantha, data= BCI_sub, corr= corGaus(form= ~ x + y, nugget = T))
plot(Variogram(modallGnug, form= ~ x + y))


#Try rational quadratic
modallR = gls(Drypetes.standleyi ~ Cordia.lasiocalyx + Hirtella.triandra + Picramnia.latifolia + Quassia.amara + Tabernaemontana.arborea + Trattinnickia.aspera + Xylopia.macrantha, data= BCI_sub, corr= corRatio(form= ~ x + y))
plot(Variogram(modallR, form= ~ x + y))

#Try rational quadratic with nugget
modallRnug = gls(Drypetes.standleyi ~ Cordia.lasiocalyx + Hirtella.triandra + Picramnia.latifolia + Quassia.amara + Tabernaemontana.arborea + Trattinnickia.aspera + Xylopia.macrantha, data= BCI_sub, corr= corRatio(form=~x + y, nugget = T))
plot(Variogram(modallRnug, form=~x + y))


#Anova of all models
anova(modall, modall1, modall1nug, modallG, modallGnug, modallR, modallRnug, test = F)

```
Including exponential error in the model changed the coefficients of each variable, decreasing some while increasing others. The exponential error model had the best fit according to the ANOVA. Including a nugget or trying other types of error did not improve the fit. As with the first model, there seems to be systematic error in the variogram including exponential error. However, other types of error did not provide a better fit.

While adding error improved both models, the improvement was much more noticeable in the model with only one explanatory variable. This is most likely because there is no apparent correlation between D. standleyi and the other species as a whole, but there is between D. standleyi and T. aspera alone.